The Two Dualities of Computation: Negative and Fractional Types

### 2012-03-10 rational.tex
_

\begin{abstract}
  Every functional programmer knows about sum and product types, {{a+b}} and
  {{a*b}} respectively. Negative and fractional types, {{a-b}} and {{a/b}}
  respectively, are much less known and their computational interpretation is
  unfamiliar and often complicated. We show that in a programming model in
  which information is preserved (such as the model introduced in our recent
  paper on \emph{Information Effects}), these types have particularly natural
  computational interpretations. Intuitively, values of negative types are
  values that flow ``backwards'' to satisfy demands and values of fractional
  types are values that impose constraints on their context.  The combination
  of these negative and fractional types enables greater flexibility in
  programming by breaking global invariants into local ones that can be
  autonomously satisfied by a subcomputation. Theoretically, these types give
  rise to \emph{two} function spaces and to \emph{two} notions of
  continuations, suggesting that the previously observed duality of
  computation conflated two orthogonal notions: an additive duality that
  corresponds to backtracking and a multiplicative duality that corresponds
  to unification.
\end{abstract}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Constructions in the Additive Monoid}

>>>>

\begin{abstract}
  Every functional programmer knows about sum and product types, {{a+b}} and
  {{a*b}} respectively. Negative and fractional types, {{a-b}} and {{a/b}}
  respectively, are much less known and their computational interpretation is
  unfamiliar and often complicated. We show that in a programming model in
  which information is preserved (such as the model introduced in our recent
  paper on \emph{Information Effects}), these types have particularly natural
  computational interpretations. Intuitively, values of negative types are
  values that flow ``backwards'' to satisfy demands and values of fractional
  types are values that impose constraints on their context.  The combination
  of these negative and fractional types enables greater flexibility in
  programming by breaking global invariants into local ones that can be
  autonomously satisfied by a subcomputation. Theoretically, these types give
  rise to \emph{two} function spaces and to \emph{two} notions of
  continuations, suggesting that the previously observed duality of
  computation conflated two orthogonal notions: an additive duality that
  corresponds to backtracking and a multiplicative duality that corresponds
  to constraint propagation.
\end{abstract}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Constructions in the Additive Monoid}
\label{sec:monoidal-constructions}
